### 2015-12-10 paper/4-performance.tex
Introduction to parallel simulation performance

\begin{figure}[h]
	\includegraphics[width=\columnwidth]{fig/fig1.eps}
	\caption{Queue benchmark results}
	\label{fig:Queue_benchmark}
\end{figure}

@@

%TODO only include sequential results?
\begin{figure}[h]
	\includegraphics[width=.5\textwidth]{fig/fig3.eps}
	\label{fig3.eps}
	\caption{Interconnect}
	\label{fig:Interconnect}
\end{figure}

\subsection{Parallel Simulation Execution Time}
%TODO explain that we will now compare different synchronization protocols
%TODO compare different synchronization protocols with adevs' conservative, showing that we are faster (or at least, not that much slower) than adevs
By default, the benchmarks use 4 kernels for parallel simulation.
\subsubsection{Devstone}
The flattened models are allocated to kernels by giving each kernel a distinct section of the chain, resulting in a low ratio of inter-kernel to intra-kernel messages. For the optimistic case, this can cause more reverts since the kernels will start to drift faster as the model-count increases. Furthermore, optimistic synchronization is quite sensitive to an increase in kernels, since the delay before a revert propagates, increases. This benchmark requires a specific warm-up time; for $n=d\times w$ models, it takes $\textup{timeadvance}()*(n-1)$ transitions to activate the last model in the chain. When using parallel processing, this can be reduced to $\textup{timeadvance}()*(n\frac{kernels-1}{kernels} -1)$ before the last kernel becomes active.\\
Figure~\ref{fig:DevstoneParallel} shows that dxex outperformance adevs (in both optimistic and conservative case).
\begin{figure}
	\includegraphics[width=.5\textwidth]{fig/fig2.eps}
	\label{fig:DevstoneParallel}
	\caption{DevStone parallel}
\end{figure}
\subsubsection{PHold}
%Phold
In Phold, the allocation is specified in the benchmark itself. Each kernel manages a single node with a constant set of sub-nodes. The parameter $R$ (which is here set to $10$) determines the percentage of remote destination models.\\
The dynamic dependency graph is a very sparse version of the static dependency graph, penalizing the conservative case. The lookahead is $\epsilon$, so the conservative case spends most of its time crawling in steps of $\epsilon$. Since the dependency graph between kernels is a complete graph, this is not a simulation that scales in our implementation. For $N$ kernels, each kernel has to query the null-time of $N-1$ kernels, resulting in O($N^2$) polling behaviour. In a non-cyclic simulation with a non-trivial lookahead (like in, e.g., Devstone), that choice does pay off (see Figure~\ref{fig:DevstoneParallel}). Adevs's lesser performance is due in part to their lookahead management, which after profiling shows to spend a non-trivial amount of time in exception handling code.\\
The optimistic case suffers little from the above problems; due to the high interconnectivity, however, a cascading revert is still possible. With the percentage of remotes equal to $100$, Phold reflects interconnect in behaviour, which is why the $R$ value is not dimensioned here. A revert is very expensive in PHold due to our usage of C++11's random nr generators. The cost of a revert is dominated by the recalculation of destination models, not in allocating/deallocating states/messages. Once a revert happens the drift between kernels increases fast, increasing the likelihood of more reverts. Despite all this, optimistic can for low $R$ values quickly exploit the uncertainty that slows down conservative in this benchmark.

>>>>

\begin{figure}
	\includegraphics[width=\columnwidth]{fig/fig1.eps}
	\caption{Queue benchmark results}
	\label{fig:Queue_benchmark}
\end{figure}

@@

%TODO only include sequential results?
\begin{figure}
	\includegraphics[width=\columnwidth]{fig/fig3.eps}
	\label{fig3.eps}
	\caption{Interconnect}
	\label{fig:Interconnect}
\end{figure}

\subsection{Parallel Simulation Execution Time}
We now perform an analysis of parallel simulation execution times of our previously defined benchmarks.
For dxex, we mention results for both conservative and optimistic synchronization.
Since adevs supports only conservative synchronization, we don't mention optimistic synchronization results there.
All experiments were performed using four simulation nodes, and executed on a quad-core machine.

We highlight two main results:
(1) dxex conservative synchronization is competitive with adevs;
(2) dxex optimistic synchronization is sometimes more efficient than conservative synchronization.
This shows that our contribution, offering both conservative and optimistic synchronization, is indeed beneficial for a general-purpose simulation tools.

\subsubsection{Queue}
In the Queue model, we allocate the chain of models such that each node is responsible for a series of connected models.
This minimizes the number of inter-node messages.
As the model is a queue, however, the last models will only activate much later on in the simulation.
Since these are allocated to seperate nodes, some nodes will remain idle until simulation has progressed sufficiently far.

Similar to the sequential benchmarks, Figure~\ref{fig:queue_benchmark_parallel} shows that dxex again outperforms adevs, using both optimistic and conservative synchronization.
%TODO wait for raw results
Behaviour is exactly the same as in sequential simulation, but some speedup is achieved.

\begin{figure}
	\includegraphics[width=\columnwidth]{fig/fig2.eps}
	\label{fig:queue_benchmark_parallel}
	\caption{Queue parallel simulation results using 4 nodes.}
\end{figure}

\subsubsection{PHold}
%Phold
In Phold, the allocation is specified in the benchmark itself. Each kernel manages a single node with a constant set of sub-nodes. The parameter $R$ (which is here set to $10$) determines the percentage of remote destination models.\\
The dynamic dependency graph is a very sparse version of the static dependency graph, penalizing the conservative case. The lookahead is $\epsilon$, so the conservative case spends most of its time crawling in steps of $\epsilon$. Since the dependency graph between kernels is a complete graph, this is not a simulation that scales in our implementation. For $N$ kernels, each kernel has to query the null-time of $N-1$ kernels, resulting in O($N^2$) polling behaviour. In a non-cyclic simulation with a non-trivial lookahead (like in, e.g., Devstone), that choice does pay off (see Figure~\ref{fig:DevstoneParallel}). Adevs's lesser performance is due in part to their lookahead management, which after profiling shows to spend a non-trivial amount of time in exception handling code.\\
The optimistic case suffers little from the above problems; due to the high interconnectivity, however, a cascading revert is still possible. With the percentage of remotes equal to $100$, Phold reflects interconnect in behaviour, which is why the $R$ value is not dimensioned here. A revert is very expensive in PHold due to our usage of C++11's random nr generators. The cost of a revert is dominated by the recalculation of destination models, not in allocating/deallocating states/messages. Once a revert happens the drift between kernels increases fast, increasing the likelihood of more reverts. Despite all this, optimistic can for low $R$ values quickly exploit the uncertainty that slows down conservative in this benchmark.
